### 2013-10-18 Rcode/ash.test.rmd
added ash.test.rmd

```{r, echo=FALSE}
#Note, this needs the ashr library installed.
#To build the library run the following (from the Rcode subdirectory):
#require(devtools)
#install("../package/ashr")
#Then RESTART R before knitting the following code
```

### Project Goals

The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\hat\beta=(\hat\beta_1,\dots, \hat\beta_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).

Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements $j=1,\dots,J$ to improve inference for each individual $\beta_j$. By improved inference, we mean both
improved average accuracy of point estimates, which is
the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty.

By ``adaptive" shrinkage we 
have two key properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified. Second, the amount of shrinkage undergone by each $\hat\beta_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.

Given that shrinkage estimation is widely recognized as a powerful tool, there are suprisingly few software packages for performing the simplest type of shrinkage estimation considered here. (There are more packages for the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
The only package we have found that provides anything similar
to the functionality provided here is mixfdr (Muralidharan). Compared with mixfdr, the key features of ashr are that it
i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symettric).
NOTE: should emphasise these differences in the examples.

As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero. Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
and then applying a generic procedure, such as that of Benjamini 
and Hochberg (1995?) or Storey (2001?), designed to control or
estimate the false discovery rate (FDR) or the positive FDR (Storey, 2001?). In essence we aim to provide analagous
generic methods that work directly with two numbers for each 
measurement $(\hat\beta_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic). Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the 
uncertainty in each measurement $\hat\beta_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.

The potential for shrinkage-based estimation to
address the multiple comparisons setting 
has been highlighted
previously, including Greenland and Robins (1991),
Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]

### Methods Outline

The methods are based on treating the vectors $\hat\beta$
and $s$ as ``observed data", and then performing inference for $\beta$ from these observed data, using a standard hierarchical modelling framework
to combine information across $j=1,\dots,J$.

Specifically, 
we assume that the true 
$\beta_j$ values are independent
and identically distributed from some distribution $g(\cdot ; \pi)$, where $\pi$ is a hyper-parameter to be estimated. 
Then, given $\beta$, we assume that $(\hat\beta_j,s_j)$ are independent across $j$, and depend only on $\beta_j$. Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
\begin{align}
p(\hat\beta, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\hat\beta_j, s_j | \beta_j) \\
& = \prod_j g(\beta_j ; \pi) L(\beta_j; \hat\beta_j,s_j).
\end{align}
The specific choices of $g$ and $L$ are described below.

We fit this hierarchical model using the following "Empirical Bayes" approach. First we estimate the hyper-parameters $\pi$ by maximizing the likelihood
$$L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.$$
Then, given this estimate $\hat\pi$, we compute the conditional distributions $$p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).$$ 
In principle we would
prefer to take a full Bayes approach that accounts for uncertainty in $\pi$, but, at least for now, we compromise this principle for the simplicity of the EB approach.
[Note: a Variational Bayes version of this is
also implemented, and may become our preferred approach
after testing]

[put picture of hierarchical model here]

The conditional distributions $p(\beta_j | \hat\pi, \hat\beta, s)$ 
encapsulate uncertainty in the values for $\beta_j$, combining information across
$j=1,\dots,J$. The combining of the information occurs through estimation of
$\pi$, which involves all of the data, and it is

These conditional distributions can be conveniently summarized
in various ways, including point estimates (e.g. the posterior means or medians),
and credible intervals/regions.

The key components of this hierarchical model
are the distribution $g$ and the likelihood $L(\beta_j; \hat\beta_j, s_j)$. We make the following choices for these.

1. The likelihood for $\beta_j$ is normal, centered on $\hat\beta_j$, with standard deviation $s_j$.
That is, 
$$L(\beta_j; \hat\beta_j, s_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2]. \quad (**)$$

2. The distribution $g(\cdot; \pi)$ is a mixture of zero-centered normal distributions, 
$$g(\cdot; \pi) = \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma^2_k).$$
In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small (possibly 0), to very large --  sufficiently large that typically $\hat\pi_K=0$.

The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways. For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$. Now, if we are willing to assume that
$s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then
$L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$,
and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$,
as is often asymptotically the case, then we obtain the likelihood (**) above.

An alternative motivation is to think of this as a normal
approximation to the likelihood from the 
raw data $D_j$ that were used to compute
$\hat\beta_j$ and $s_j$. Then if we observed these data
the likelihood for $\beta$ would be 
$p(D_j | \beta_j)$, and a Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]

Using a mixture of normal distributions for $g$ 
also seems very natural: mixtures of normals provide a flexible family of distributions able to provide a good approximation to any true underlying $g$; and 
when combined with the normal likelihood they give
an analytic form for the conditional distribution $p(\beta_j | \hat\pi, \hat\beta_j, s_j)$ (also a mixture of normals).
The constraint that these normals be centered at zero may seem initially less natural. 
Certainly this constraint could be relaxed in principle.
However, we view it as a convenient way to impose an assumption 
that $g$ is unimodal with its mode at 0,
which we view as a plausible assumption in many 
settings, and one which may be helpful to avoid
"overfitting" of $g$. (Using normal distributions centered at 0
also imposes an assumption
that $g$ is symmetric about zero, which we view as less
plausible, but represents a compromise between simplicity
and flexibility.
In cases where this assumption seems wildly inappropriate one could perhaps
improve results by applying the
model separately to positive and negative values of $\hat\beta$.)

Finally, using a large number of normal components with a wide range of variances,
rather than, say, a smaller number of components with the variances
to be estimated, is simply for computational convenience. With fixed
variances there exists a very simple EM algorithm to maximise the likelihood
in $\pi$. Obtaining maximum likelihood estimates for the variances could certainly
be implemented with a little
more work, but it is unclear whether this would result in practically-important gains in many situations of interest.

### The False Sign Rate

The usual definition of the local False Discovery Rate (lfdr) for observation $j$ is
$\text{lfdr}_j = p(\beta_j = 0 | \hat\beta, s)$.
The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$), so lfdr 
gives the probability that we make a mistake if we reject
$H_j$, that is the probability that, if we reject $H_j$,
it is a  "false discovery".

As pointed out by Gelman et al, there are settings where
$\beta_j=0$ is implausible, in which case the lfdr is not
a useful concept: if every $\beta_j$ is non-zero then there
is no such thing as a false discovery and the lfdr will be 0. Gelman et al suggest that in such settings we
might replace the concept of a false discovery with the
concept of an "error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
$\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
fundamentally more interesting and useful than identifying which are non-zero. However, in such settings this usually comes down to something similar in practice: the observations that one can tell are non-zero are also the ones whose
sign can be reliably inferred. In this sense the
local false sign rate is a natural generalization of
the local false discovery rate.

Following these ideas of Gelman et al, we define the local False Sign Rate (lfsr) as 
$$\text{lfsr}_j = \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].$$
To give the intuition behind this, suppose for concreteness
that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then, if we were to declare that $\beta_j$ is negative, the probability that we have
made an error in sign would be 0.05.

Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ include the event $\beta_j=0$.
Thus, lfsr gives an upper bound for the lfdr,
and so can be used as a conservative estimate of the lfdr
if an lfdr is desired. (This may be helpful when comparing
with methods that compute an lfdr.)
In addition, we intuitively expect the difference $\text{lfsr}_j - \text{lfdr}_j$ to be small, at least when $\text{lfdr}_j$ is small, because when we are confident that a particular effect is non-zero we would expect to be similarly confident about the sign of the effect.

### Computation Outline

As outlined above, we fit the model using the following Empirical Bayes procedure:
1. Estimate $\pi$ by maximizing the likelihood $L(\pi)$.
2. Compute the conditional distributions $p(\beta_j | \hat\beta_j, s_j, \hat\pi)$.

Using a normal likelihood $L(\beta_j)$, and assuming
$g$ to be a mixture of normal distributions with fixed variances, 
yields a simple EM algorithm
for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
distributions in Step 2.

### Example 1: A simple simulated example to get started

Load in some functions.
```{r}
#setwd("~/Documents/git/ash/Rcode/")
set.seed(32327)

## load ashr library 
require("ashr")
require("qvalue")
```

```{r}
#simulate n beta-hat values, nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata=function(n, nnull, altmean, altsd, betasd){
  null = c(rep(1,nnull),rep(0,n-nnull))
  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
  betahat = rnorm(n,beta,betasd)
  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
}
```

Simulate 10000 $\beta_j$ values, 8000 from a mixture of a point mass at 0 and
2000 from $\beta \sim N(0,sd=2)$.

@@

beta.ash=ash(ss$betahat,ss$betasd)
xmin = min(ss$beta)
xmax = max(ss$beta)
x=seq(xmin,xmax,length=1000)
plot(sort(ss$beta),(1:length(ss$beta))/length(ss$beta),main="cdf of ss$beta, with fitted f overlaid",xlab="beta",ylab="cdf")
lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)

Compare with mixfdr. Note that I tried to do smaller than 1000, but mixfdr returned
errors. (Maybe this example is not so useful for comparing fdrs (as opposed to shrinkage), since the methods tend to agree where the fdr is small, which is what matters?)
```{r}
ss = simdata(1000,500,0,2,1)

beta.ash=ash(ss$betahat,ss$betasd)
beta.ash.u = ash(ss$betahat,ss$betasd, mixcompdist="uniform")
beta.ash.hu = ash(ss$betahat,ss$betasd, mixcompdist="halfuniform")

require(mixfdr)
beta.mfdr.J3 = mixFdr(ss$betahat,J=3,noiseSD=1,theonull=TRUE)
beta.mfdr.J5 = mixFdr(ss$betahat,J=5,noiseSD=1,theonull=TRUE)
beta.mfdr.J7 = mixFdr(ss$betahat,J=7,noiseSD=1,theonull=TRUE)
beta.mfdr.J9 = mixFdr(ss$betahat,J=9,noiseSD=1,theonull=TRUE)
beta.mfdr.J15 = mixFdr(ss$betahat,J=15,noiseSD=1,theonull=TRUE)
mean((beta.mfdr.J3$effectSize-ss$beta)^2)
mean((beta.mfdr.J5$effectSize-ss$beta)^2)
mean((beta.mfdr.J7$effectSize-ss$beta)^2)
mean((beta.mfdr.J9$effectSize-ss$beta)^2)
mean((beta.mfdr.J15$effectSize-ss$beta)^2)

mean((beta.ash$PosteriorMean-ss$beta)^2)
mean((beta.ash.u$PosteriorMean-ss$beta)^2)
mean((beta.ash.hu$PosteriorMean-ss$beta)^2)

#This code mostly borrowed from the examples in mixFdr in R help
m=beta.mfdr.J15
mTrue = list(pi=c(0.8,0.2),mu=c(0,0),sig=c(1,2))
par(mfcol=c(1,1))
s = seq(-5,5,by=0.01)
trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.002) # this tells it which indices to consider null
 plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
 lines(s, fdrCurve, col = 2)
points(ss$betahat,beta.ash$localfdr,col=3)
points(ss$betahat,beta.ash.u$localfdr,col=4)
points(ss$betahat,beta.ash.hu$localfdr,col=5)

Of course, if we do something where alternatives are not centered on 0 then
ash does worse!
```{r}
#This is the example from the mixfdr help example
z = c(rnorm(1000), rnorm(100,3,1))
 m = mixFdr(z, J = 3, noiseSD = 1)
 s = seq(-5,5,by=0.01)
 effCurve = effectSize(s,m)[,1]
 fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.2) # this tells it which indices to consider null

# compare to the Bayes estimator for this problem
 mTrue = list(pi = c(10,1)/11, mu = c(0,3), sig = c(1,1))
# note that the model parameters are in terms of the MARGINAL not of the prior
 trueEff = effectSize(s,mTrue)[,1]
 trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
 par(mfrow = c(2,1))
 plot(s, trueEff, t = 'l', main = "Effect Size - Black is true", xlab = "z", ylab = "E(delta|z)", ylim = c(-3,3))
 lines(s, effCurve, col = 2)
 plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))

z.ash = ash(z,1,mixcompdist="halfuniform")
points(z,z.ash$localfdr,col=3)
```

Checking coverage
```{r}

Things to emphasise for paper:
- the number of components is not critical; the unimodal constraint is enough. You can increase the number of components arbitrarily and the likelihood remains bounded.
- the ability to incorporate item-specific measurement error
- the ability to compare a model in which effects are proportional to error
with model in which effects

- unimodal distributions could be argued to be a primary motivation for shrinkage

- bayesian coverage intervals (attempt to) give much stronger guarantess than standard confidence intervals. Eg among the 0.95 intervals excluding 0,
less than 5% should actually be 0?

Some advantages of the unimodal constraint:
-more statsitically efficient (if true); show with small sample sizes.
- less sensitive to number of components? Can allow number to tend to infinity?
- may be less sensitive to local optima?
could demonstrate this by looking at convergence more carefully, and comparing
results with random starting points.

- it allows us to easily just vary sigma on a grid, and fit pi, which
makes allowing different noise levels really easy!

## The number of components is not critical

OR - this section could be more generally about selecting between
models (normal, uniforms etc) using log-likelihood.

Because of the uni-modal constraint, the number of mixture components
is generally not critical, at least provided it is ``sufficiently large". 
Indeed, as the number of components tends to infinity, the likelihood
is bounded above, and even for large numbers of components and small amounts of data the inferred underlying distributions tend not to be too crazy, showing few signs of ``overfitting".

However, it is true that the normal distributions tend to yield smoother estimated underlying distributions. Similarly, using the posterior mean for $\pi$, rather
than the maximum likelihood estimate, tends to lead to somewhat smoother fitted distributions. In addition, with small amounts of data the underlying
distributions are inevitably not well determined by the data, and the fits may vary depending on the underlying assumptions made (e.g. number of components or distribution of the components); however, even then shrinkage-based estimates
of the betas can be relatively robust.

The following example illustrates this.
We take 100 observations from 0.5N(-1,1)+0.5N(1,1)
and add N(0,1) noise.
```{r}
set.seed(200)
sampsize=50
beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
betahat = beta+rnorm(2*sampsize)

beta.ash.default = ash(betahat,1) #run with default parameters for comparison
beta.ash.vb.default = ash(betahat,1,VB=TRUE) #run with default parameters for comparison

beta.ash.100 = ash(betahat,1,mixsd = seq(0.1,5,length=100))
beta.ash.100u = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=100))
beta.ash.100hu = ash(betahat,1,mixcompdist="halfuniform",mixsd = seq(0.1,5,length=100))

beta.ash.vb.100 = ash(betahat,1,mixsd = seq(0.1,5,length=100),VB=TRUE)
beta.ash.vb.100u = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=100),VB=TRUE)
beta.ash.vb.100hu = ash(betahat,1,mixcompdist="halfuniform",mixsd = seq(0.1,5,length=100),VB=TRUE)

x=seq(-4,4,length=1000)
plot(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1,ylim=c(0,0.45),main="Fitted densities vs truth",type="l",ylab="density")
lines(x,density(beta.ash.default,x),col=2)
lines(x,density(beta.ash.100,x),col=2)
lines(x,density(beta.ash.100u,x),col=3)
lines(x,density(beta.ash.100hu,x),col=4)

plot(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1,ylim=c(0,0.45),main="Fitted densities (VB) vs truth",type="l",ylab="density")
lines(x,density(beta.ash.vb.default,x),col=2)
lines(x,density(beta.ash.vb.100,x),col=2,lty=2)
lines(x,density(beta.ash.vb.100u,x),col=3,lty=2)
lines(x,density(beta.ash.vb.100hu,x),col=4,lty=2)

plot(x,0.5*pnorm(x,-1,1)+0.5*pnorm(x,1,1),col=1,main="Fitted cdfs vs truth",type="l",ylab="density")
lines(x,cdf.ash(beta.ash.100,x),col=2,lty=2)
lines(x,cdf.ash(beta.ash.100u,x),col=3,lty=2)
lines(x,cdf.ash(beta.ash.100hu,x),col=4,lty=2)

plot(x,0.5*pnorm(x,-1,1)+0.5*pnorm(x,1,1),col=1,main="Fitted cdfs (VB) vs truth",type="l",ylab="density")
lines(x,cdf.ash(beta.ash.vb.default,x),col=2,lty=2)
lines(x,cdf.ash(beta.ash.vb.100,x),col=2,lty=2)
lines(x,cdf.ash(beta.ash.vb.100u,x),col=3,lty=2)
lines(x,cdf.ash(beta.ash.vb.100hu,x),col=4,lty=2)

c(loglik.ash(beta.ash.100,betahat,1),get_loglik(beta.ash.100u),get_loglik(beta.ash.100hu))
c(loglik.ash(beta.ash.vb.100,betahat,1),loglik.ash(beta.ash.vb.100u,betahat,1),loglik.ash(beta.ash.vb.100hu,betahat,1))

rmse=function(a,b){sqrt(mean((a$PosteriorMean-b)^2))}
c(rmse(beta.ash.default,beta),rmse(beta.ash.100,beta),rmse(beta.ash.100u,beta),rmse(beta.ash.100hu,beta),rmse(beta.ash.vb.default,beta),rmse(beta.ash.vb.100,beta),rmse(beta.ash.vb.100u,beta),rmse(beta.ash.vb.100hu,beta))

plot(beta.ash.default$PosteriorMean,beta.ash.100u$PosteriorMean,col=3,cex=0.4,xlim=c(-4,4),ylim=c(-4,4))
points(beta.ash.default$PosteriorMean,beta.ash.100hu$PosteriorMean,col=4,cex=0.4)
abline(a=0,b=1)

plot(betahat,ifelse(pnorm(betahat)<0.5,pnorm(betahat),1-pnorm(betahat)))
points(betahat,beta.ash.default$localfsr,col=2)
points(betahat,beta.ash.100u$localfsr,col=3)
points(betahat,beta.ash.100hu$localfsr,col=4)

so it seems that the localfsr is almost the same as for a "flat" prior $g$.
That is, in this case, the underlying distribution is flat relative to the
error terms, and so there is little shrinkage effect except at the tails
(where the fsr is basically 0 anyway).

```{r}
sampsize=5000
beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
betahat = beta+rnorm(2*sampsize)

beta.ash = ash(betahat,1)
beta.ash.u=list()
for(i in 1:10){
  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=10*i))
}

plot(beta.ash,-4,4,col=2)
x=seq(-4,4,length=1000)
lines(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1)
for(i in 1:4){
lines(x,density(beta.ash.u[[i]],x),col=2+i)
}

plot(unlist(lapply(beta.ash.u,ashr:::get_loglik)))

beta.ash=ash(ss$betahat,ss$betasd,auto=FALSE)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#these to test the VB version 
beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,VB=TRUE, prior="uniform")
beta.ash.vb.nullbiased = ash(ss$betahat, ss$betasd,VB=TRUE, prior="nullbiased")

@@

@@

It seems that in this case the ash q values underestimate the
FDR slightly. Possibly this is the assymetry.
Try shrinking positive and negatives separately:
```{r}
pos= betahat>0
betapos.ash=ash(betahat[pos],s[pos])
betaneg.ash = ash(betahat[!pos],s[!pos])
lfdr = rep(0,length(betahat))
lfdr[pos] = betapos.ash$localfdr
lfdr[!pos] = betaneg.ash$localfdr
qv = qval.from.localfdr(lfdr)
o = order(qv)
plot(cumsum(truenull[o])/(1:10000),qv[o],type="l")
abline(a=0,b=1)
```
No, that's not it.

So maybe it is the non-monotonic behaviour of the true beta values in this case.
Let's try putting in a component at -1.5 to make the data fit this assumption better.

@@

### Motivation

It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values. The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
In genomics, for example, a very common goal
is to compare the mean
expression (activity) level of many genes in two conditions.
Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
measurements are made on only a small number of
samples in each condition - sometimes as few as one
sample in each condition. Thus the error in estimates
of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
in estimates of $\beta_j$ still greater.

A fundamental idea is
that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.

Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by Greenland and Robins. It isn't the number of tests that is relevant (the false
discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.

Performing multiple comparisons, or multiple tests, is often
regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.

Imagine that you are preparing to perform 1 million tests,
each based on a $Z$ score that is assumed to be $N(0,1)$ under the null.
You first order these tests randomly, and begin by 
performing the first test, which returns a $Z$ score of 4.
At this point you are interrupted by a friend, who asks how the analysis is going. "It's early days, but looking promising" you reply. Well, who wouldn't? If the aim is to find lots of significant differences, a strong first result is surely a good outcome.

At this point you have reason to expect that many of the
subsequent tests also output strong results.

Now consider two alternative scenarios for the remaining 999,999 tests. In the first scenario, the remaining tests
produce $Z$ values that fit very well with the null, closely following a standard normal distribution; in the second
scenario a large proportion of the remaining tests, say 50 percent, show outcomes that lie outside of $[-4,4]$.

If your friend enquired after your analysis again, your
response would surely differ in the first scenario ("Oh, it didn't pan out so well after all") vs the second ("It went great"). Further,
in the first scenario, if your friend pressed you further about the results of the first test,
you would likely, I think, be inclined to put them down to chance. In contrast, in the second scenario, the first test turned out to be, as you hoped, a harbinger of good things to come, and in this scenario you would likely regard that test as likely corresponding to a true discovery.

The key point is that it is the *outcomes* of the tests, not the *number* of tests, that impacts interpretation of that first test.

(Some may by pondering whether the fact that you are about to perform another
999,999 tests should be considered pertinent in responding to your friend. Our view is that it is  irrelevant.
The standard frequentist framework would disagree, because
it requires the analyst to consider hypothetical repetitions
of the "experiment", and so the fact that the experiment
consists of a million tests is pertinent. However, this
argument is a distraction from the main point.)

Indeed, we believe that the practice of focussing on the *number* of tests performed is

Focussing on the number of tests performed can be
seen as an approximation.

The standard argument is that,
when performing multiple tests, some will be significant
just by chance.

>>>>

```{r}
require("ashr")
require("qvalue")
require("testthat")

set.seed(100)

#simulate n beta-hat values, nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata=function(n, nnull, altmean, altsd, betasd){
  null = c(rep(1,nnull),rep(0,n-nnull))
  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
  betahat = rnorm(n,beta,betasd)
  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
}

ss = simdata(10000,8000,0,2,1)

set.seed(133)
system.time((beta.ash=ash(ss$betahat,ss$betasd,randomstart=TRUE)))
set.seed(133)
system.time((beta.ash2=oldash(ss$betahat,ss$betasd,auto=TRUE,randomstart=TRUE)))

expect_that(beta.ash$PosteriorMean, equals(beta.ash2$PosteriorMean))
expect_that(beta.ash$qvalue, equals(beta.ash2$qvalue))
expect_that(beta.ash$PositiveProb, equals(beta.ash2$PositiveProb))

system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.uniform2 = oldash(ss$betahat, ss$betasd, VB=TRUE, auto=TRUE,prior="uniform")))
expect_that(beta.ash.vb.uniform$PosteriorMean, equals(beta.ash.vb.uniform2$PosteriorMean))
expect_that(beta.ash.vb.uniform$qvalue, equals(beta.ash.vb.uniform2$qvalue))
expect_that(beta.ash.vb.uniform$PositiveProb, equals(beta.ash.vb.uniform2$PositiveProb))

system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
system.time((beta.ash.vb.null2 = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
expect_that(beta.ash.vb.null$PosteriorMean, equals(beta.ash.vb.null2$PosteriorMean))
expect_that(beta.ash.vb.null$qvalue, equals(beta.ash.vb.null2$qvalue))
expect_that(beta.ash.vb.null$PositiveProb, equals(beta.ash.vb.null2$PositiveProb))

```{r}
require("ashr")
require("qvalue")

set.seed(100)

#Test mixVBEM
abf = rbind(c(1,0,0,0),c(0,1,0,0),c(0,0,1,0),c(0,0,1,0),c(0,0,1,0))
eps = 1e-10
abf[abf==0] = eps #replace 0 with small number
print(all.equal(mixVBEM(abf,c(1,1,1,1))$post,c(2,2,4,1)))
print(all.equal(mixVBEM(abf,c(1,2,1,1))$post,c(2,3,4,1)))

#simulate n beta-hat values, nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata=function(n, nnull, altmean, altsd, betasd){
  null = c(rep(1,nnull),rep(0,n-nnull))
  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
  betahat = rnorm(n,beta,betasd)
  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
}

ss = simdata(10000,8000,0,2,1)

system.time((beta.ash=ash(ss$betahat,ss$betasd,sigmaavec=c(0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24))))
system.time((beta.ash.auto = ash(ss$betahat, ss$betasd)))
system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior="strongnull")))

hist(ss$beta,prob=TRUE,breaks=seq(floor(min(ss$beta)),ceiling(max(ss$beta)),length=20))
x= seq(-4,4,length=10000)
lines(x,density(beta.ash,x),col=2)
lines(x,density(beta.ash.auto,x),col=3)
lines(x,density(beta.ash.vb.uniform,x),col=4)
lines(x,density(beta.ash.vb.null,x),col=5)
summary(beta.ash)
summary(beta.ash.auto)
summary(beta.ash.vb.uniform)
summary(beta.ash.vb.null)

ss = simdata(10000,10000,0,2,1)
system.time((beta.ash=ash(ss$betahat,ss$betasd)))
system.time((beta.ash.auto = ash(ss$betahat, ss$betasd)))
system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))

hist(ss$beta,prob=TRUE,breaks=seq(floor(min(ss$beta)),ceiling(max(ss$beta)),length=20))
x= seq(-4,4,length=10000)
lines(x,density(beta.ash,x),col=2)
lines(x,density(beta.ash.auto,x),col=3)
lines(x,density(beta.ash.vb.uniform,x),col=4)
lines(x,density(beta.ash.vb.null,x),col=5)
beta.ash$fitted.f
beta.ash.auto$fitted.f
beta.ash.vb.uniform$fitted.f
beta.ash.vb.null$fitted.f

beta.ash.vb.uniform$fit$loglik
beta.ash.vb.uniform$fit$converged
```

@@

beta.ash=ash(ss$betahat,ss$betasd,sigmaavec=c(0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24))
beta.ash.auto = ash(ss$betahat, ss$betasd)
#these to test the VB version 
beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")
beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)

@@

Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
```{r}
summary(beta.ash)
summary(beta.ash.auto)
summary(beta.ash.vb.uniform)
summary(beta.ash.vb.null)
summary(beta.ash.pm)
```

@@

It seems that in this case the ash q values underestimate the
FDR slightly. Possibly this is the assymetry.
Try shrinking positive and negatives separately:
```{r}
pos= betahat>0
betapos.ash=ash(betahat[pos],s[pos])
betaneg.ash = ash(betahat[!pos],s[!pos])
lfdr = rep(0,length(betahat))
lfdr[pos] = betapos.ash$localfdr
lfdr[!pos] = betaneg.ash$localfdr
qv = qval.from.localfdr(lfdr)
o = order(qv)
plot(cumsum(truenull[o])/(1:10000),qv[o],type="l")
abline(a=0,b=1)
```

@@
